Relationships in the extremes and their influence on competition and coexistence

### 2021-08-02 paper/Paper.Rnw
more climate chnage refs and lit review

%***DAN: stopped here 2021 07 20
[As we know], anthropogenic climate change threatens almost all aspects of life including via extreme climatic events, which have only become more extreme - increasing in magnitude and frequency - and have contributed to both the increase in global means as well as an increase in environmental variability (citation). 
There are many studies on the influence of environmental extremes and variability on competition and coexistence \citep{Adler_2006,  Parepa_2013, Diez_2012}; however, the study of the influence of relationships and interactions between drivers in their extremes on competition and coexistence, is lacking in the scientific literature.
%;we review these studies here  In one of the first empirical tests for storage effect and its requirements, \cite{Adler_2006} showed that climate variability - including extreme events - promoted the coexistence of grass species in Kansas prairie, specifically their ability to recover from low densities. Moreover, in studies on the invasibility of plants, the density and competitive balance of invasive Japanese knotweed increases as resource and thus climate variability increased.\citep{Parepa_2013} 
Interacting environmental drivers can be asymmetrically related in the extreme values (as shown in \cite{Gosh_2020_AER}). 
Environmental drivers directly influence competitive pressure experienced by populations through environmental response; competition is always a function of environmental response \citep{Chesson_2012, Ellner_2016}.
Thus we have implemented our tools for the study of ATAs to explore the importance of environmental relationships in their extremes on competition and coexistence. 
Environmental variability has been classically shown to promote coexistence via storage mechanisms in a temporal lottery model of competition \citep{Chesson_1981, Warner_1985}, hence we continue/move forward with this simple model to examine the effect of extreme relationships between environmental responses on the contribution of storage effects to coexistence. Moreover, the lottery model within the framework of coexistence theory provides a suitable opportunity to explore relationships in their extremes and the consequences to coexistence.

>>>>

%***DAN: stopped here 2021 07 20
%[As we know], anthropogenic climate change threatens almost all aspects of life including via extreme climatic events, which have only become more extreme - increasing in magnitude and frequency - and have contributed to both the increase in global means as well as an increase in environmental variability (citation). 
%There are many studies on the influence of environmental extremes and variability on competition and coexistence \citep{Adler_2006,  Parepa_2013, Diez_2012}; however, the study of the influence of relationships and interactions between drivers in their extremes on competition and coexistence, is lacking in the scientific literature.
%;we review these studies here  In one of the first empirical tests for storage effect and its requirements, \cite{Adler_2006} showed that climate variability - including extreme events - promoted the coexistence of grass species in Kansas prairie, specifically their ability to recover from low densities. Moreover, in studies on the invasibility of plants, the density and competitive balance of invasive Japanese knotweed increases as resource and thus climate variability increased.\citep{Parepa_2013} 
Environmental fluctuations are becoming more influenced by extreme climatic events (Hansen et al 2012) as extreme events increase in frequency, duration, and intensity \citep{Meehl_2004}. Extreme events (heat-waves, floods, hurricanes, etc.) alter the means and variation of environmental fluctuations, and thus have implications for coexistence and competition. Competitive output and community composition of grassland plants has been shown to be impacted by heat-waves and extreme flooding \citep{White_2000, White_2001, Hoover_2014}. Studies on the direct effect of extreme events on competing species are slim (but see above); however, the consequences of change in environmental means, which may be due to extreme events or gradual climate change in general, on coexistence is more commonly studied.  \cite{Descombes_ 2020}, through experimental manipulation, found that range shifts, as a product of climate change, favored coexistence between alpine plants while \cite{Usinowicz_2021} predict increases in competitive pressures as the climate change induced range shifts lead populations to interact with new conspecifics. Similarly, \cite{diPaola_2013} predict the collapse of coexistence between Mediterranean tree species as rainfall levels are forecasted to decrease. Extreme events constitute large disturbances, quantitatively amounting to large fluctuations in environmental drivers (citation). More variable environmental fluctuations are even less so considered compared to changes in spatial means; although, environmental variability has been analytically shown to be important for classical coexistence \citep{Chesson_1981, Warner_1985}. Moreover, increased variability has been found to promote coexistence in plants communities \citep{Adler_2006,  Angert_2009}, but disrupt coexistence in birds \citep{Stenseth_2002, Stenseth_2015}.

Interacting environmental drivers can be asymmetrically related in the extreme values (as shown in \cite{Gosh_2020_AER}). 
Environmental drivers directly influence competitive pressure experienced by populations through environmental response; competition is always a function of environmental response \citep{Chesson_2012, Ellner_2016}.
Thus we have implemented our tools for the study of ATAs to explore the importance of environmental relationships in their extremes on competition and coexistence. 
Environmental variability has been classically shown to promote coexistence via storage mechanisms in a temporal lottery model of competition \citep{Chesson_1981, Warner_1985}, hence we continue/move forward with this simple model to examine the effect of extreme relationships between environmental responses on the contribution of storage effects to coexistence. Moreover, the lottery model within the framework of coexistence theory provides a suitable opportunity to explore relationships in their extremes and the consequences to coexistence.
